---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fable.prophet
[![Travis build status](https://travis-ci.org/mitchelloharawild/fable.prophet.svg?branch=master)](https://travis-ci.org/mitchelloharawild/fable.prophet)
[![Codecov test coverage](https://codecov.io/gh/mitchelloharawild/fable.prophet/branch/master/graph/badge.svg)](https://codecov.io/gh/mitchelloharawild/fable.prophet?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

This package provides a tidy R interface to the prophet forecasting procedure using [fable](https://github.com/tidyverts/fable). This package makes use of the [prophet package](https://cran.r-project.org/package=prophet) for R.

## Installation

You can install the development version of fable.prophet from [Github](https://github.com/mitchelloharawild/fable.prophet) with:

``` r
# install.packages("remotes")
remotes::install_github("mitchelloharawild/fable.prophet")
```

## Example

Suppose we wanted to model Australia's monthly turnover for cafes, restaurants and catering services. The data is available from the Australian Bureau of Statistics catalogue 8501.0, and in the [tsibbledata](https://github.com/tidyverts/tsibbledata) package.

```{r data, message=FALSE}
library(tsibble)
library(dplyr)
cafe <- tsibbledata::aus_retail %>% 
  filter(Industry == "Cafes, restaurants and catering services")
```
```{r plot, echo = FALSE}
library(ggplot2)
cafe %>% 
  ggplot(aes(x = Month, y = Turnover, colour = State)) + 
  geom_line() + 
  ylab("Turnover (millions $AUD)")
```

Each series generally exhibits an increasing trend with an annual seasonal pattern that varies proportionally to the level of the series. At a monthly level, any holiday effects can be modelled using a seasonal term. A piecewise linear trend is included by default, and so it is not included in the model specification below.

```{r model}
library(fable.prophet)
fit <- cafe %>% 
  model(
    prophet = prophet(Turnover ~ season("year", 4, type = "multiplicative"))
  )
```

```{r mable}
fit
```

The above output confirms that this Prophet model has been fitted to each of the time series. Components from this model can be extracted:

```{r components}
components(fit)
```
```{r components-plot, echo = FALSE, message = FALSE}
library(lubridate)
components(fit) %>%
  ggplot(aes(x = Month, y = trend, colour = State)) + 
  geom_line()
components(fit) %>%
  ggplot(aes(x = month(Month), y = year, 
             colour = State, group = interaction(year(Month), State))) + 
  geom_line() + 
  scale_x_continuous(breaks = 1:12, labels = month.abb) + 
  xlab("Month")
```

Note that the annual seasonal pattern does not change very quickly, although it does differ slightly between years. A very differently seasonal pattern can be seen for the Northern Territory. We can also produce forecasts for each of these series over the next two years.

```{r forecast}
fc <- fit %>% 
  forecast(h = 24)
```

```{r fable, echo = FALSE}
fc
cafe %>% 
  ggplot(aes(x = Month, y = Turnover, colour = State)) + 
  geom_line() + 
  autolayer(fc)
```

